A Dissection of Foregrounds in Epoch of Reionization Observations from the Murchison Widefield Array

### 2015-01-06 ms.tex
Restructured section 2 and minor changes addressing many referee comments

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous {\bf pathfinders and} precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{tin13,bow13,lon09}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}. {\bf The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER}.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}.

@@

Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies on the sky as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
\begin{align}\label{eqn:obsvis}
  V_{bf} &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. We wish to {\bf note} that this equation is valid {\bf even for wide--field applications where the sky cannot be approximated to be flat}.

We define the {\it delay spectrum}, $V_{b\tau}\equiv V_{b\tau}(\boldsymbol{b},\tau)$, to be the inverse Fourier transform of $V_{bf}$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  V_{b\tau} &\equiv \int V_{bf}\,W'_f\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W'_f\equiv W'_f(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}

Equation~\ref{eqn:obsvis} can be {\bf equivalently re-expressed} as:
\begin{align}\label{eqn:vis}
  V_{uf} &\equiv V_{uf}(\boldsymbol{u},f) \nonumber\\
  &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
\end{align}
where, $\boldsymbol{u}$ denotes the transverse spatial frequency vector. $\boldsymbol{u}$ is directly related to $\boldsymbol{b}$ as $\boldsymbol{u}\equiv \boldsymbol{b}/\lambda$, and to the transverse spatial wavenumber mode, $\boldsymbol{k}_\perp$, as $\boldsymbol{k}_\perp\equiv 2\pi\boldsymbol{u}/D(z)$, where, $D(z)$ is the transverse comoving distance at redshift $z$. Hence, $V_{uf}=V_{uf}(\boldsymbol{b}/\lambda,f)=V_{bf}(\lambda\boldsymbol{u},f)$. It is important to note that for a given antenna pair, $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.

Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight. $\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time. It is directly related to the line--of--sight wavenumber, $k_\parallel\approx 2\pi\eta\,f_{21}H_0\,E(z)/(c(1+z)^2)$, where $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible. Thus,
\begin{align}\label{eqn:los-transform}
  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
\end{align}
represents the true spatial Fourier representation of the three--dimensional sky brightness distribution. This representation has been discussed in detail in \citet{mor04}. The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by $P(\boldsymbol{k}_\perp,k_\parallel)\propto |V_{u\eta}|^2$.

While $V_{b\tau}$ and $V_{u\eta}$ in equations~\ref{eqn:obsvis} and \ref{eqn:vis} resemble each other very closely, they are not quite identical. The difference basically arises from the fact noted above that $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not. \citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.

Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity. Foregrounds can be described in both the delay spectrum and Fourier frameworks. For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite their subtle differences.

In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power $|V_{u\eta}|^2$ in Fourier modes containing critical information about spatial scales. For purposes of illustration later in this paper, we often use the quantity, $|V_{b\tau}|$.

Note that since {\it visibility} is a complex quantity, its Fourier transform is Nyquist sampled over positive and negative delays. Negative and positive delay\footnote{The terms delay and lag are used interchangeably in this paper.} correspond to the two hemispheres of the sky as transected by the antenna spacing vector. In this paper, $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward directions at the telescope site. Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively. Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward.

Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction, form the $x$--axis. $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries are called horizon delay limits and are shown by solid lines. The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis. This effect has been quantified in \citet{thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
\includegraphics[width=\linewidth]{f1.eps}
\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented. The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber). The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber). Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz. The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside. Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits. Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the instrumental transfer function are expected to be relatively free of foreground contamination. Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper. \label{fig:fourier-space}}
\end{figure}

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67\fdg 5 \leq \theta_b < 112\fdg 5$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.

@@

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$.

The MWA passband of width 30.72~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter. The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$. After correcting for the shape of these coarse channels, the fine channels (40~kHz width) at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level. Any spectral flags, or other known weights constitutes $W_f$ in equation~\ref{eqn:obsvis}.

Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation. Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation. In our work, we have chosen a $T_\textrm{sys}$ that matches the thermal noise observed in the data.

@@

The MWA targets two primary low-foreground fields for reionization observations. Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30$\arcdeg. The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile. The pointing system is capable of steering to points on a regular $\sim$7\arcdeg~ grid. During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary (once in $\sim 30$~minutes). This process is repeated throughout the course of the observation $\approx 4.86$~hours. Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.

@@

The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at $-1.92$~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}. It has a flux density limit of 25~mJy and a declination range of -12\arcdeg~ to -40\arcdeg~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by small reflections in subset of cables.

@@

The {\it off--zenith} pointing has emission notably higher than in the {\it zenith} pointing inside the {\it foreground wedge}. This is shown later to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky. Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings. Emission in the {\it zenith} pointing is more centrally concentrated due to phasing of power--pattern to the zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith but is also responsive to emission from angles far away from zenith.

>>>>

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous {\bf pathfinders and} precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}. {\bf The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER}.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}. {\bf Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity}.

@@

Interferometer array data known as {\it visibilities}, $V_{bf}\equiv V_{bf}(\boldsymbol{b},f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power--pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_f\equiv W_f(f)$ denotes instrumental bandpass weights, then $V_{bf}$ can be written as (with a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}):
\begin{align}\label{eqn:obsvis}
  V_{bf} &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. We wish to {\bf note} that this equation is valid {\bf even for wide--field applications where the sky cannot be approximated to be flat}.

{\bf The {\it delay spectrum}, $V_{b\tau}\equiv V_{b\tau}(\boldsymbol{b},\tau)$, is defined as} the inverse Fourier transform of $V_{bf}$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  V_{b\tau} &\equiv \int V_{bf}\,W'_f\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W'_f\equiv W'_f(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}
$V_{b\tau}$ is expressed in observer's units as Jy~Hz.

\subsection{Resemblance to Spatial Power Spectrum}\label{sec:power-spectrum}

Equation~\ref{eqn:obsvis} can be {\bf equivalently re-expressed} as:
\begin{align}\label{eqn:vis}
  V_{uf} &\equiv V_{uf}(\boldsymbol{u},f) \nonumber\\
  &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_f\,e^{-i2\pi\boldsymbol{u}\cdot\hat{\boldsymbol{s}}}\,\dif\Omega,
\end{align}
where, $\hat{\boldsymbol{s}}$ is measured with reference to a location on the sky referred to as the {\it phase center}, and $\boldsymbol{u}\equiv (u,v,w)$ denotes the spatial frequency vector. $w$ is aligned parallel to the direction of the phase center, while $u$ and $v$ lie on the transverse plane perpendicular to it. For measurements that are constrained to lie on this plane, we can choose $w=0$ without loss of generality and $\boldsymbol{u}$ effectively reduces to $\boldsymbol{u}\equiv (u,v)$, a two--dimensional vector. Then, $\boldsymbol{u}$ is directly related to the transverse spatial wavenumber mode as: 
\begin{align}\label{eqn:k-perp}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi\boldsymbol{u}}{D}, 
\end{align}
where, $D\equiv D(z)$ is the transverse comoving distance at redshift $z$. % Hence, $V_{uf}=V_{uf}(\boldsymbol{b}/\lambda,f)=V_{bf}(\lambda\boldsymbol{u},f)$. It is important to note that for a given antenna pair, $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not.

Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight. $\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time. It is directly related to the line--of--sight wavenumber, 
\begin{align}\label{eqn:k-prll}
  k_\parallel &\approx \frac{2\pi\eta\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible. Thus,
\begin{align}\label{eqn:los-transform}
  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
\end{align}
represents the true spatial Fourier representation of the three--dimensional sky brightness distribution. This approach has been discussed in detail in \citet{mor04}. The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by \citep{mor04,mcq06,par12a}: 
\begin{align}\label{eqn:true-power_spectrum}
  P(\boldsymbol{k}_\perp,k_\parallel) &\simeq |V_{u\eta}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant. Thus, $V_{u\eta}$ inferred from observations, also in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$.

Without loss of generality, the phase center can be assumed to be the zenith relative to the local tangent plane. Then $\boldsymbol{u}$ lies on this plane for measurements constrained to be on it. If the array of antennas are also coplanar lying on the local tangent plane, then $\boldsymbol{u}=\boldsymbol{b}/\lambda$. Under such circumstances, equations~\ref{eqn:obsvis} and \ref{eqn:delay} closely resemble equations~\ref{eqn:vis} and \ref{eqn:los-transform} respectively. However, they are not quite identical to each other. The difference basically arises from the fact noted above that $\boldsymbol{b}$ is independent of frequency while $\boldsymbol{u}$ is not. \citet{par12b} and \citet{liu14a} have discussed the mathematical correspondence between the two.

% Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. 
Foregrounds can be described in both the delay spectrum and Fourier frameworks. For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences. In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of aforementioned cosmological H{\sc i} power spectrum derived, we define:
\begin{align}\label{eqn:fake-power-spectrum}
  P'(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
where, $\boldsymbol{u}$ and $\eta$ in equations~\ref{eqn:k-perp} and \ref{eqn:k-prll} have been correspondingly replaced with $\boldsymbol{b}/\lambda$ and $\tau$ respectively.

In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales. For purposes of illustration later in this paper, we often use the quantities $|V_{b\tau}|$ and $P'(\boldsymbol{k}_\perp,k_\parallel)$.

Note that since {\it visibility} is a complex quantity, its {\bf delay} transform {\bf contains} positive and negative delays. Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector. In this paper, $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site. Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at positive and negative delays, respectively. Similarly, an object in the northern sky will appear at a positive delay at an antenna spacing oriented northward. For all observations used in this study, we use the zenith as the phase center.

{\bf We give a brief overview of some parameters of Fourier space which is generic to all experiments that use a similar approach}. Figure~\ref{fig:fourier-space} illustrates the Fourier space, {\bf specific to the MWA}, in which the delay (and power) spectra of H{\sc i} from the EoR are presented in this paper. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction {\bf (tangent plane to the celestial sphere)}, form the $x$--axis. $\tau\approx\eta\propto k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines. The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis \citep{par12b,thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. However, the instrument used in our study, namely the MWA, has a passband constructed using coarse channels with a characteristic shape giving rise to periodic gratings thereby resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/fourier_space_185.0_MHz_30.7_MHz.eps}
\includegraphics[width=\linewidth]{f1.eps}
\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are presented. The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber). The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber). Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz. The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside. Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits. Regions excluding the {\it wedge} and narrow extensions of the {\it wedge} (white unshaded strips) caused by convolution with the {\bf instrument's spectral transfer function} are expected to be relatively free of foreground contamination. Due to an undesirable grating pattern (dot--dashed lines) of the instrument used in this study, namely the MWA, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the {\it EoR window} in this paper. \label{fig:fourier-space}}
\end{figure}

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.

@@

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. In its final configuration the MWA is a 128--tile interferometer array capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. The array is arranged as a centrally condensed core of $\sim$~300~m --- there are many spacings in the range 5--50~m --- and a radial density that falls off as the inverse of the radius, with the longest baselines at 3~km \citep{bea12}. Here we focus on antenna spacings $|\boldsymbol{b}| \le 200$~m (spatial scales relevant to reionization) and a frequency range 170--200~MHz. The center frequency of 185~MHz corresponds to a redshift, $z\simeq 6.68$. {\bf At this frequency, we use $A_\textrm{e}=10.5$~m$^2$}.

The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter. The coarse channel shape is obtained using a 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$. {\bf Each of these coarse bands consists of 32 fine channels of width 40~kHz each}. After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level. Any spectral flags, or other known weights constitutes $W_f$ in equation~\ref{eqn:obsvis}.

% Thermal noise in simulated visibilities is estimated assuming a system temperature of $T_\textrm{sys}=95$~K per polarization for all frequency channels for all antenna pairs throughout the course of the observation. Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation. In our work, we have chosen a $T_\textrm{sys}$ that matches the thermal noise observed in the data.

@@

The MWA targets two primary low-foreground fields for reionization observations. Here, we focus on the field at RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$. The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of the dipoles in an MWA tile. The pointing system is capable of steering to points on a regular $\sim$7\arcdeg~ grid. During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary (once in $\sim 30$~minutes). This process is repeated throughout the course of the observation $\approx 4.86$~hours. Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.

@@

The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours ($-1.92$~hours in the figure)} and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}. It has a flux density limit of 25~mJy and a declination range of $-12\arcdeg$~ to $-40\arcdeg$~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by small reflections in subset of cables.

@@

The {\it off--zenith} pointing {\bf has notably higher power in the {\it foreground wedge} than in the {\it zenith} pointing}. This is shown later in \S\ref{sec:diffuse} to be due to response of the array to the bright Galactic center and Galactic plane in the westward sky. Consequently, the contamination into the {\it EoR window} is also higher, especially on short antenna spacings. Emission in the {\it zenith} pointing is more centrally concentrated {\bf due to a zenith--centered power--pattern and faster suppression of response away from zenith, whereas the {\it off--zenith} pointing has a power--pattern phased eastward of zenith and is also more responsive to emission farther away from the primary lobe relative to that in the {\it zenith} pointing. The power pattern is discussed in detail in \S\ref{sec:instrument_model} and Figure~\ref{fig:power_pattern}}.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\includegraphics[width=\linewidth]{f3.eps}
\caption{Delay spectra of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set. {\bf In both cases, the delay spectra are phased to the zenith and thus the {\it foreground wedge} boundaries are symmetric around $\tau=0$}. For reference, the $k$--space equivalent of $\tau$ and $|\boldsymbol{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline. Smooth spectrum objects predominantly reside in this {\it foreground wedge}. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:fhd_data}}
\end{figure}
